Displacement of fishing effort by Large Scale Marine Protected Areas

### 2018-09-29 rmd/Manuscript.Rmd
Updated latest analyses and manuscript

---
title: "Displacement of fishing effort by Large Scale Marine Protected Areas"
subtitle: "Updated on `r Sys.Date()`"
author: |
  Juan Carlos Villaseñor-Derbez^1^
  John Lynham^2^
abstract: "Large-scale Marine Protected Areas (LSMPAs) have seen a significant increase over the last years. Fishing effort is effectively eliminated within these protected areas upon implementation. The benefits of reducing effort have been largely studied, and include increases in abundance, biomass, and diversity within the bounded regions. These no-take zones may produce spillover effects, which provide fish for outside areas. However, the economic and ecological implications of displacing fishing effort are not yet fully understood. Novel data products that track fishing effort at the vessel-level allow us to identify changes in fleet- and vessel-level behavior upon the implementation of protected areas, as well as how these redistribute. This papers evaluates the implications of implementing LSMPA, by evaluating changes in fishing hours, showing that vessels in the effected region reduce fishing effort after the implementation of PIPA. Our results are robust to a set of specifications. We also track the relative spatial allocation of fishing events thorugh time, and identify that areas closer to PIPA show an increase in relative fishing hourse due to the displacement of PIPA-fishing vessels. Our results not only provide an impact evaluation of the effect of LSMPAs on fishing activity, but provide insights into vessel redistribution dynamics, which may have ecological and economic implications."
date: |
  ^1^Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, CA
  ^2^Department of Economics, University of Hawaii at Manoa, Honolulu, HI
output:
  bookdown::pdf_document2:
    toc: yes
    toc_depth: 4
    citation_package: natbib
    keep_tex: yes
    number_sections: no
header-includes:
  - \usepackage{float}
  - \floatplacement{figure,table}{H}
  - \usepackage{lineno}
  - \linenumbers
bibliography: references.bib
biblio-style: nature
---

@@

This section is divided into two main parts. First, we provide a general description of AIS data and the process of identification of vessel-level events done by Global Fishing Watch [^1]. Alongside, we describe the subset of data that we use for these analyses. When relevant, we also point out possible shortcommings in the data, or factors that must be considered in the later analyses. Then we move on to explain our identification strategy, and the main analyses that we undertake.

[^1]: [Global Fishing Watch](globalfishingwatch.org)

@@

Automatic Identification Systems are on-board devices intended to provide at-sea saftey and prevent ship collisions by broadcasting vessel locations to surrounding vessels. These broadcasted positions can be recorded by satellites and land-based antenas. GFW uses a neural network to infer vessel characteristics and whether each broadcasted position represents a fishing event, thus allowing us to estimate near real-time fishing events globally \citep{kroodsma_2018}. The recent addition of satellites that can receive AIS signals causes an apparent increase in the number of broadcasted AIS messages (*i.e.* points) and therefore fishing hours. It is therefore important to be cautious when comparing temporal trends. The variability in AIS data and dynamic ocean conditions require that temporal trends be taken into account. We do that by incorporating a series of controls, which are defined in the following section.

@@

```{r}
effort_by_vessel <- read.csv(file = here::here("data", "effort_by_vessel.csv"))
```

```{r}
# Just to get pipa vessels that fished inside pipa before 2015
pipa_vessels <- effort_by_vessel %>% 
  filter(treated) %$%
  unique(mmsi) %>% 
  length()

n_prev <- effort_by_vessel %>% 
  filter(treated, !post) %$%
  unique(mmsi) %>% 
  length()

n_post <- effort_by_vessel %>% 
  filter(treated, post, baci_strict) %$%
  unique(mmsi) %>% 
  length()
```

The GFW data contain `r pipa_vessels` purse seiners and longliners that have fished within PIPA waters. From these, only `r n_prev` did so at least once before 2015, but only `r n_post` continued to fish after 2015[^2]. Vessels that fished within PIPA before implementation might stop fishing afterwards, therefore not being observable in the post-treatment period. New vessels might have also entered the fishery after PIPA closure, and were likely not exposed to the policy intervention in the pre-treatment period. Therefore, we define our treatment and control groups as follows. 
The treatment group contains all vessels that fished within PIPA at least once before the closure, and that continued to fish elsewhere afterwards. The control group has vessels that never fished within PIPA waters, belong to other PNA countries, and have fished in surrounding areas (*i.e.* PNA-countries' EEZ) before and after PIPA closure. Figure \ref{fig:baci_strict} provides a visual representation of the vessel-level that make up each of group through time.

Tables \ref{tab:baci_n}, \ref{tab:baci_n_s}, and \ref{tab:baci_h} show the number of vessels, number of vessels following a BACI design, and fishing hours, respectively. Tables show data grouped by gear (longlines, purse seines), group (treated, control), and period (before, after). The relative change (After / Before) is also shown. Table \ref{tab:baci_n} shows information for all fishing vessels in the dataset, while Tables \ref{tab:baci_n_s} and \ref{tab:baci_h} show information for vessels that meet the BACI design (i.e. excludes vessels that only appear at one point in time).

![\label{fig:baci_strict}Stream of fishing events by vessels (stacked on the y-axis due to space) through time (x-axis). Each dot represent a vessel-month, with colors indicating the pre and post periods, and pannels separating between gear and treated or control groups.](../img/BACI_strict_gear.png)

[^2]: The `r n_prev - n_post` missing vessels might have exited the fishery, been decomissioned or sold (therefore chaning their AIS and mmsi), or turned off their AIS transmiters. On either case, we are not able to observe these.

```{r}
effort_by_vessel %>% 
  # filter(baci_strict) %>%
  group_by(gear, treated, post, mmsi) %>% 
  count() %>% 
  ungroup() %>% 
  group_by(gear, treated, post) %>% 
  summarize(boats = n()) %>% 
  spread(post, boats) %>% 
  mutate(change = `TRUE` / `FALSE`) %>% 
  knitr::kable(col.names = c("Gear", "Treatment", "Before", "After", "Change (A / B)"),
               caption = "\\label{tab:baci_n}Number of fishing vessels (identified by mmsi) for each gear before and after PIPA implementation.",
               booktabs = T) %>% 
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

```{r}
effort_by_vessel %>% 
  filter(baci_strict) %>%
  group_by(gear, treated, post, mmsi) %>% 
  count() %>% 
  ungroup() %>% 
  group_by(gear, treated, post) %>% 
  summarize(boats = n()) %>% 
  spread(post, boats) %>% 
  mutate(change = `TRUE` / `FALSE`) %>% 
  knitr::kable(col.names = c("Gear", "Treatment", "Before", "After", "Change (A / B)"),
               caption = "\\label{tab:baci_n_s}Number of fishing vessels (identified by mmsi) for each gear before and after PIPA implementation.",
               booktabs = T) %>% 
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

```{r}
effort_by_vessel %>%
  filter(baci_strict) %>%
  group_by(gear, treated, post) %>% 
  summarize(hours = mean(hours, na.rm = T)) %>% 
  spread(post, hours) %>% 
  mutate(change = `TRUE` / `FALSE`) %>% 
  knitr::kable(col.names = c("Gear", "Treatment", "Before", "After", "Change (A / B)"),
               caption = "\\label{tab:baci_h}Mean fishing hours for each gear before and after",
               booktabs = T) %>% 
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

@@

### Change in behavior

@@

$$
y_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \mu_t + \phi_t + \gamma_i + \epsilon_{i,t}
$$

Where $y_{ijkl}$ is the variable of interest for vessel $i$ in time period $t$. A dummy variable $Post_i$ takes the value of 0 for all dates prior to PIPA implementation and a value of 1 por all dates including and following PIPA implementation. $Treat_i$ is a dummy variable indicating whether a vessel belongs to the control ($Treat_i = 0$) or treatment ($Treat_i = 1$) group. $\alpha$ is the standard intercept, $\beta_1$ captures the *ex post* change, $\beta_2$ captures the difference between treated and control groups, and $\beta_3$ is our parameter of interest: de DiD estimate capturing the treatment effect. Finally, $\mu_t$, $\phi_t$, and $\gamma_i$ represent month-, year-, and flag-level dummies that account for seasonality or country-level management interventions.

@@

### Vessel displacement and redistribution

```{r}
displacement_data <- readRDS(file = here::here("data", "displacement_data.rds"))
```

Our second part of the analyses focuses on the redistribution of fishing effort. In other words, identifying where do vessels that used to fish inside PIPA go after its establishment. We calculate the monthly relative distribution of fishing hours by all treated vessels across all fished EEZs and the high seas. These trends are shown in Figure \ref{fig:redist_trend}. EEZs that had sproadic fishing events were pooled into a group of "others", leaving us with a total of n = `r length(unique(displacement_data$country))` spatially defined regions (*i.e.* EEZs, High Seas, "other EEZs"). Figure \ref{fig:redist_trend_spa} shows the yearly average of the relative spatial allocation of fishing efforts; this is the spatial representation of \ref{fig:redist_trend} averaged by year.

@@

```{r, fig.cap = "\\label{fig:redist_trend}"}
displacement_data %>% 
  ggplot(aes(x = date, y = h_prop, group = post)) +
  geom_smooth(method = "lm") +
  geom_point() +
  geom_vline(xintercept = lubridate::date(paste(2015, 1, 1, sep = "/")), linetype = "dashed") +
  scale_fill_brewer(palette = "Paired") +
  ggtheme_plot() +
  facet_wrap(~country, scales = "free_y")
```

@@

<!-- Meter tabla ba_disp aqui -->

Fig. \ref{fig:all_vessels} shows that mean fishing hours for purse seiners have an abrupt increase, just before January 1st, 2015. This trend is observed for both treated and control groups. Longliners, however, show a more stable trend. The number of mmsi codes increases through time. This can largel be explained by the addition of more satellits, which increasedetectability of vessels. As expected total fishing hours follow a similar trend to that of mmsi numbers. Results of the regressions for each gear type are shown in Tables \ref{tab:purse} and \ref{tab:long}. Column (1) presents the DiD regression with no fixed effects, column (2) includes month fixed effects, column (3) includes month and year fixed effects, and column (4) includes month, year, and flag fixed effects. Figures \ref{fig:puse} and \ref{fig:long} represent the same coefficient estimates.

```{r, echo = F, fig.cap = "\\label{fig:all_vessels}Fishing hours and number of vessels by month for all vessels. Vertical dashed line indicates PIPA closure."}
effort_by_vessel %>% 
  filter(baci_strict) %>%
  group_by(year, gear, treated, post, month) %>% 
  summarize(total_hours = sum(hours),
            mean_hours = mean(hours),
            n_mmsi = n()) %>% 
  ungroup() %>% 
  mutate(date = lubridate::date(paste(year, month, "01", sep = "/"))) %>% 
  select(-year) %>% 
  gather(variable, value, -c(date, month, post, gear, treated)) %>% 
  ggplot(aes(x = date, y = value, color = treated)) +
  geom_line() +
  geom_point() +
  facet_wrap(gear~variable, scales = "free", ncol = 3) +
  theme_bw() +
  scale_color_brewer(palette = "Set1") +
  geom_vline(xintercept = lubridate::date(paste(2015, 01, "01", sep = "/")), linetype = "dashed")
```

>>>>

---
title: "Displacement of fishing effort by Large Scale Marine Protected Areas"
subtitle: "Updated on `r Sys.Date()`"
author: |
  Juan Carlos Villaseñor-Derbez^1^
  John Lynham^2^
abstract: "Large-scale Marine Protected Areas (LSMPAs) have seen a significant increase over the last years. Fishing effort is effectively eliminated within these protected areas upon implementation. The benefits of reducing effort have been largely studied, and include increases in abundance, biomass, and diversity within the bounded regions. These no-take zones may produce spillover effects, which provide fish for outside areas. However, the economic and ecological implications of displacing fishing effort are not yet fully understood. Novel data products that track fishing effort at the vessel-level allow us to identify changes in fleet- and vessel-level behavior upon the implementation of protected areas, as well as how these redistribute. This papers evaluates the implications of implementing LSMPA, by evaluating changes in fishing hours, showing that vessels in the effected region reduce fishing effort after the implementation of PIPA. Our results are robust to a set of specifications. We also track the relative spatial allocation of fishing events thorugh time, and identify that areas closer to PIPA show an increase in relative fishing hourse due to the displacement of PIPA-fishing vessels. Our results not only provide an impact evaluation of the effect of LSMPAs on fishing activity, but provide insights into vessel redistribution dynamics, which may have ecological and economic implications."
date: "^1^Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, CA\\newline^2^Department of Economics, University of Hawaii at Manoa, Honolulu, HI"
output:
  bookdown::pdf_document2:
    toc: no
    citation_package: natbib
    keep_tex: yes
    number_sections: no
fontsize: 12pt
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \usepackage{lineno}
  - \linenumbers
bibliography: references.bib
# biblio-style: nature
---

@@

This section is divided into two main parts. First, we provide a general description of AIS data and the process of identification of vessel-level events done by Global Fishing Watch[^1]. Alongside, we describe the subset of data that we use for these analyses. When relevant, we also point out possible shortcommings in the data, or factors that must be considered in the later analyses. Then we move on to explain our identification strategy, and the main analyses that we undertake.

[^1]: Global Fishing Watch: [globalfishingwatch.org](globalfishingwatch.org)

@@

Automatic Identification Systems are on-board devices intended to provide at-sea saftey and prevent ship collisions by broadcasting vessel locations to surrounding vessels. These broadcasted positions can be recorded by satellites and land-based antenas. GFW uses a neural network to infer vessel characteristics and whether each broadcasted position represents a fishing event, thus allowing us to estimate near real-time fishing events globally since 2012 \citep{kroodsma_2018}. The recent addition of satellites that can receive AIS signals causes an apparent increase in the number of broadcasted AIS messages (*i.e.* points) and therefore fishing hours. The variability in AIS data and ocean conditions require that temporal trends be taken into account. We do that by incorporating a series of controls, which are defined in the following section.

@@

```{r}
effort_by_vessel <- readRDS(file = here::here("data", "effort_by_vessel.rds")) %>% 
  filter(year < 2018)
```

```{r}
# Just to get pipa vessels that fished inside pipa before 2015
pipa_vessels <- effort_by_vessel %>% 
  filter(treated) %$%
  unique(mmsi) %>% 
  length()
```

Our data contain over 45 million individual AIS messages for 371 purse seiners and longliners. A total of 233 vessels have fished within PIPA waters; 217 did so at least once before 2015. However, not all vessels continued to fish elsewhere after PIPA implementation: 34 vessels have no recorded AIS messages after 2015[^2], leaving us with `r pipa_vessels` vessels that fished inside PIPA before its implementation, and continued to fish elsewhere afterward. Vessels that fished within PIPA before implementation might stop fishing afterwards, therefore not being observable in the post-treatment period. New vessels might have also entered the fishery after PIPA closure, and were likely not exposed to the policy intervention in the pre-treatment period. Therefore, we define our treatment and control groups as follows.

The treatment group contains all vessels (n = `r pipa_vessels`) that fished within PIPA at least once before the closure, and that continued to fish elsewhere afterwards. Vessels in the control group meet al three of the following conditions: i) Vessels never fished within PIPA waters, ii) vessels belong to other PNA countries, and iii) vessels have fished in surrounding areas (*i.e.* PNA-countries' EEZ) before and after PIPA closure. For each vessel meeting these characteristics, we calculate their total monthly fishing hours. Figure \ref{fig:baci_strict} provides a visual representation of the vessel-level streams of fishing events that make up each group through time. Tables \ref{tab:baci_n_s} and \ref{tab:baci_h} show the number of vessels following a BACI design, and fishing hours, respectively. Tables show data grouped by gear (longlines, purse seines), group (treated, control), and period (before, after). The relative change in fishing hours (After / Before) is also shown.

<!-- **Do we know what is causing this? [first sentence in fist paragraph of results] I'm guessing this can't be Blue Paradox stuff because we see it for the control group as well who are supposed to have never fished within PIPA.** -->

<!-- **Is the larger number of longliners surprising? I thought PIPA was known as a purse seining ground and less so for longlining.** -->

[^2]: The `r n_prev - n_post` missing vessels might have exited the fishery, been decomissioned or sold (therefore changing their AIS and mmsi), or turned off their AIS transmiters. In either case, we are not able to observe these.

```{r, fig.width = 8, fig.height = 6, fig.cap = "\\label{fig:baci_strict}Stream of fishing events by vessels through time. Each line represents a vessel, with dots indicating months with fishing activity and colors indicating the pre and post periods. Pannels separating between gear and treated or control groups."}
filter(effort_by_vessel, baci_strict) %>% 
  mutate(gear = ifelse(gear == "purse_seines", "Purse seiners", "Drifting longliners")) %>% 
  group_by(mmsi, date, gear, post, treated) %>%
  summarize(h = sum(hours, na.rm = T)) %>%
  ungroup() %>% 
  arrange(treated, post, mmsi) %>%
  mutate(vessel = group_indices(., paste(gear, treated, mmsi)),
         treated = ifelse(treated, "Treated", "Control")) %>% 
  ggplot(aes(
    x = date,
    y = vessel,
    fill = post,
    color = post,
    group = paste(vessel, post)
    )) +
  geom_point(size = 1) +
  theme_bw() +
  facet_wrap(gear ~ treated, scales = "free_y") +
  ggtheme_plot() +
  scale_color_brewer(palette = "Set1") +
  scale_y_continuous(breaks = NULL, labels = NULL) +
  theme(legend.position = "none")
```

```{r}
effort_by_vessel %>% 
  filter(baci_strict) %>%
  group_by(gear, treated, mmsi) %>% 
  count() %>% 
  ungroup() %>% 
  group_by(gear, treated) %>% 
  summarize(boats = n()) %>% 
  knitr::kable(col.names = c("Gear", "Treatment", "n"),
               caption = "\\label{tab:baci_n_s}Number of fishing vessels (identified by mmsi) by gear and treatment group.",
               booktabs = T) %>% 
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

```{r}
effort_by_vessel %>%
  filter(baci_strict) %>%
  group_by(gear, treated, post) %>% 
  summarize(hours = mean(hours, na.rm = T)) %>% 
  spread(post, hours) %>% 
  mutate(change = `TRUE` / `FALSE`) %>% 
  knitr::kable(col.names = c("Gear", "Treatment", "Before", "After", "Change (A / B)"),
               caption = "\\label{tab:baci_h}Mean fishing hours and relative change by gear and treatment group.",
               booktabs = T) %>% 
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

@@

@@

$$
y_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \mu_1Y_t + \mu_2Y_t^2 + \phi_t + \gamma_i + \epsilon_{i,t}
$$

Where $y_{i,t}$ is the variable of interest for vessel $i$ in time period $t$. A dummy variable $Post_i$ takes the value of 0 for all dates prior to PIPA implementation and a value of 1 por all dates including and following PIPA implementation. $Treat_i$ is a dummy variable indicating whether a vessel belongs to the control ($Treat_i = 0$) or treatment ($Treat_i = 1$) group. $\alpha$ is the standard intercept, $\beta_1$ captures the *ex post* change, $\beta_2$ captures the difference between treated and control groups, and $\beta_3$ is our parameter of interest: de DiD estimate capturing the treatment effect. Finally, $\mu_1$ and $\mu_2$ are coefficients for a second order polynomial for years ($Y_t$)[^4], while $\phi_t$ and $\gamma_i$ represent month-, and flag-level dummies that account for seasonality or country-level management interventions.

@@

[^4]: An earlier specification included years as a dummy variable. Such results are included in the appendx.

```{r}
displacement_data <- readRDS(file = here::here("data", "displacement_data.rds")) %>% 
  filter(year < 2018)
```

Our second part of the analyses focuses on the redistribution of fishing effort. In other words, identifying where do vessels that used to fish inside PIPA go after its establishment. We calculate the monthly relative distribution of fishing hours by all treated vessels across all fished EEZs and the high seas. These trends are shown in Figures \ref{fig:redist_trend_ps} and \ref{fig:redist_trend_ll}, and the relative change is presented in Table \ref{tab:ba_disp}. EEZs that had sproadic fishing events were pooled into a group of "others", leaving us with a total of n = `r length(unique(displacement_data$country[displacement_data$gear == "purse_seines"]))` and n = `r length(unique(displacement_data$country[!displacement_data$gear == "purse_seines"]))` spatially defined regions (*i.e.* EEZs, High Seas, "other EEZs") for purse seiners and longliners, respectively.

<!-- Like before, we include a dummy variable for months and allow for a second degree polynomial trend through time. -->

All regression coefficients were estimatedd via ordinary least squares, and heteroskedastic-robust standard errors were calculated. All analyses were performed in R version `r paste(version$major, version$minor, sep = ".")`) \citep{rcore_2018}. Raw data and code used in this work are available on [github](https://github.com/jcvdav/MPA_displacement).

<!-- Figure \ref{fig:redist_trend_spa} shows the yearly average of the relative spatial allocation of fishing efforts; this is the spatial representation of \ref{fig:redist_trend} averaged by year. -->

@@

```{r}
ba_disp_ps <- displacement_data %>% 
  filter(gear == "purse_seines") %>% 
  group_by(country, post) %>% 
  summarize(h_prop_m = mean(h_prop)) %>% 
  spread(post, h_prop_m) %>% 
  mutate(change = (`TRUE` - `FALSE`) * 100) %>% 
  select(country, change)

ba_disp_ll <- displacement_data %>% 
  filter(!gear == "purse_seines") %>% 
  group_by(country, post) %>% 
  summarize(h_prop_m = mean(h_prop)) %>% 
  spread(post, h_prop_m) %>% 
  mutate(change = (`TRUE` - `FALSE`) * 100) %>% 
  select(country, change)

ba_disp_ps %>% 
  mutate(gear = "Purse seiners") %>% 
  rbind(mutate(ba_disp_ll, gear = "Longliners")) %>% 
  spread(gear, change) %>% 
  knitr::kable(digits = 2,
               booktabs = T,
               caption = "\\label{tab:ba_disp}") %>% 
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

<!-- **An additional test to add to the wish list is to see if vessels are fishing closer to each other following the PIPA closure.** -->

Our data suggest that purseiners and longliners have different responses to the implementation of a Large-Scale Marine Protected Area. Fig. \ref{fig:all_vessels} shows that mean fishing hours for purse seiners have an abrupt increase, just before January 1st, 2015. This trend is observed for both treated and control groups. Longliners, however, show a more stable trend. The number of mmsi codes increases through time. This can largely be explained by the addition of more satellites, which increase detectability of vessels. As expected, total fishing hours follow a similar trend to that of mmsi numbers.

```{r, echo = F, fig.cap = "\\label{fig:all_vessels}Fishing hours and number of vessels by month for all vessels. Vertical dashed line indicates PIPA closure."}
effort_by_vessel %>% 
  filter(baci_strict) %>%
  group_by(year, gear, treated, post, month) %>% 
  summarize(total_hours = sum(hours),
            mean_hours = mean(hours),
            n_mmsi = n()) %>% 
  ungroup() %>% 
  mutate(date = lubridate::date(paste(year, month, "01", sep = "/"))) %>% 
  select(-year) %>% 
  gather(variable, value, -c(date, month, post, gear, treated)) %>% 
  ggplot(aes(x = date, y = value, color = treated)) +
  geom_line() +
  geom_point() +
  facet_wrap(gear~variable, scales = "free", ncol = 3) +
  ggtheme_plot() +
  scale_color_brewer(palette = "Set1") +
  geom_vline(xintercept = lubridate::date(paste(2015, 01, "01", sep = "/")), linetype = "dashed")
```

Our DiD anaylsis shows that treated purse seiners reduce their fishing effort after PIPA implementation in the order of 16 hours per month. This result is robust and significant (*p < 0.05*) for all model specifications, with the effect varying between $\beta_3 = -16.457$ and $\beta_3 = -18.709$. Model specifications that include the year polynomial show lower values for the $\beta_1$ coefficient associated to the $Post_t$ policy dummy, and show positive and negative values for $\mu_1$ and $\mu_2$, the linear and cuadratic terms, respectively. These effectively represent the patterns observed in Figure {fig:all_vessels}.

Longliners show a similar pattern of effort reduction. However, the magnitude of the $\beta_3$ coefficient is smaller (ranging from  $\beta_3 = -9.851$ to $\beta_3 = -14.850$) and not significant across all model specifications. This, along with higher standard error values suggest that longliners have a smaller and more variable response to the implementation of LSMPAs.

Regressions coefficients for each gear type are shown in Tables \ref{tab:purse} and \ref{tab:long}. Column (1) presents the DiD regression with no fixed effects, column (2) includes month fixed effects, column (3) includes month and the second degree polynomial for years, and column (4) includes all of the above and country-level fixed effects.

model_data <- effort_by_vessel %>% 
  filter(baci_strict,
         gear == "purse_seines") %>% 
  mutate(post = post * 1,
         treated = treated * 1,
         year2 = year^2,
         mmsi_c = as.factor(mmsi))

m1 <- lm(hours ~ post*treated, data = model_data)
m2 <- lm(hours ~ post*treated + month_c, data = model_data)
m3 <- lm(hours ~ post*treated + month_c + year + year2, data = model_data)
m4 <- lm(hours ~ post*treated + month_c + year + year2 + flag, data = model_data)

stargazer::stargazer(models,
                     se = commarobust::makerobustseslist(models),
                     t.auto = T,
                     p.auto = T,
                     type = "latex",
                     omit = c("flag", "month"),
                     add.lines = list(c("Month FE", "No", "Yes", "Yes", "Yes"),
                                      c("Flag FE", "No", "No", "No", "Yes")),
                     omit.stat = c("adj.rsq", "f", "ser"),
                     header = F,
                     title = "\\label{tab:purse}Fishing hours from GFW for purse seiners (n = 106; 38 control, 68 treatment). Asterisks indicate significance levels. Numbers in parenthesis represent heteroskedastic-robust standard errors.")
```

```{r, fig.width = 5, fig.height = 3, fig.cap = "\\label{fig:puse}Coefficient estimates for each model. Top pannel indicates variable, x-axis represents model specification, and y-axis coefficient estimate.", eval = F}
map_df(models, broom::tidy, .id = "model") %>%
  filter(term %in% c("(Intercept)", "post", "treated", "post:treated")) %>%
  ggplot(aes(x = model, y = estimate)) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error),
                size = 1,
                width = 0,
                color = "red") +
  geom_point(color = "steelblue", size = 3) +
  facet_wrap(~term, scale = "free") +
  ggtheme_plot()
```

model_data <- effort_by_vessel %>% 
  filter(baci_strict,
         !gear == "purse_seines") %>% 
  mutate(post = post * 1,
         treated = treated * 1,
         year2 = year^2,
         mmsi_c = as.factor(mmsi))

m1 <- lm(hours ~ post*treated, data = model_data)
m2 <- lm(hours ~ post*treated + month_c, data = model_data)
m3 <- lm(hours ~ post*treated + month_c + year + year2, data = model_data)
m4 <- lm(hours ~ post*treated + month_c + year + year2 + flag, data = model_data)

stargazer::stargazer(models,
                     se = commarobust::makerobustseslist(models),
                     t.auto = T,
                     p.auto = T,
                     type = "latex",
                     omit = c("flag", "month"),
                     add.lines = list(c("Month FE", "No", "Yes", "Yes", "Yes"),
                                      c("Flag FE", "No", "No", "No", "Yes")),
                     omit.stat = c("adj.rsq", "f", "ser"),
                     header = F,
                     title = "\\label{tab:long}Fishing hours from GFW for longliners (n = 203; 88 control, 115 treatment). Asterisks indicate significance levels. Numbers in parenthesis represent heteroskedastic-robust standard errors.")
```

```{r, fig.width = 5, fig.height = 3, fig.cap = "\\label{fig:long}Coefficient estimates for each model. Top pannel indicates variable, x-axis represents model specification, and y-axis coefficient estimate.", eval = F}
map_df(models, broom::tidy, .id = "model") %>%
  filter(term %in% c("(Intercept)", "post", "treated", "post:treated")) %>%
  ggplot(aes(x = model, y = estimate)) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error),
                size = 1,
                width = 0,
                color = "red") +
  geom_point(color = "steelblue", size = 3) +
  facet_wrap(~term, scale = "free") +
  ggtheme_plot()
```

<!-- ## Effort displacement and redistribution -->

Recall that to evaluate the redistribution of fishing effort we only track fishing vessels that belong to the treated group. In this case, we calculate the prportion of fishing effort allocated every month to each spatially explicit region outlined by EEZs and the high seas. For purse seiners, these represent 9 main EEZs, PIPA, the high seas, and a group of other EEZs. Figure \ref{fig:redist_trend_ps} shows the monthly relative fishing hours that each region received by all `r pipa_vessels` treated vessels. The top-left pannel shows the change in fishing effort inside PIPA, including the preemtive fishing and immediate reduction previously reported \citep{mcdermott_2018}.

The change in the relative allocation fishing effort by purse seiners increases in eight of the 12 regions after PIPA implementation \ref{tab:disp_mod}. The largest increase is observed for the I-Kiribati EEZ, with an average increase of 0.11 (*p < 0.001*). In other words, the redistribution of treated vessels caused a 10\% increase in the *relative* allocation of fishing effort within I-Kiribati waters. The only decrease is observed for Papua New Guinea, but the coefficient is not significant. Figure \ref{fig:map_change_ps} provides a spatial representation of these changes. It is evident that the increase in relative fishing effort is greater for for regions closer to PIPA.

```{r, fig.cap = "\\label{fig:redist_trend_ps}"}
displacement_data %>% 
  filter(gear == "purse_seines") %>% 
  ggplot(aes(x = date, y = h_prop, group = post)) +
  geom_smooth(method = "lm", color = "black", size = 0.5, linetype = "dashed") +
  geom_point(size = 1) +
  geom_vline(xintercept = lubridate::date(paste(2015, 1, 1, sep = "/")), linetype = "dashed") +
  scale_fill_brewer(palette = "Paired") +
  ggtheme_plot() +
  facet_wrap(~country, scales = "free_y")
```

```{r, fig.cap = "\\label{fig:redist_trend_ll}", eval = F}
displacement_data %>% 
  filter(!gear == "purse_seines") %>% 
  ggplot(aes(x = date, y = h_prop, group = post)) +
  geom_smooth(method = "lm") +
  geom_point() +
  geom_vline(xintercept = lubridate::date(paste(2015, 1, 1, sep = "/")), linetype = "dashed") +
  scale_fill_brewer(palette = "Paired") +
  ggtheme_plot() +
  facet_wrap(~country, scales = "free_y")
```

@@

```{r, fig.cap = "\\label{fig:map_change_ps}"}
pipa_map_ba_disp <- pipa %>% 
  mutate(country = "PIPA") %>% 
  left_join(ba_disp_ps, by = "country")

# Los colores del mapa deben estar dados por el Post:Country
eez %>% 
  left_join(ba_disp_ps, by = c("ISO_Ter1" = "country")) %>%
  filter(!is.na(change)) %>% 
  ggplot() +
  geom_sf(aes(fill = change), color = "black") +
  geom_sf(data = pipa_map_ba_disp, aes(fill = change), linetype = "dashed", color = "red") +
  ggtheme_plot() +
  scale_fill_gradientn(colors = colorRamps::matlab.like(10))
```

```{r, fig.cap = "\\label{fig:map_change_ps}"}
pipa_map_ba_disp <- pipa %>% 
  mutate(country = "PIPA") %>% 
  left_join(ba_disp_ll, by = "country")

# Los colores del mapa deben estar dados por el Post:Country
eez %>% 
  left_join(ba_disp_ll, by = c("ISO_Ter1" = "country")) %>%
  filter(!is.na(change)) %>% 
  ggplot() +
  geom_sf(aes(fill = change), color = "black") +
  geom_sf(data = pipa_map_ba_disp, aes(fill = change), linetype = "dashed", color = "red") +
  ggtheme_plot() +
  scale_fill_gradientn(colors = colorRamps::matlab.like(10))
```

```{r, results = "asis"}
# Dont have this here. Just write the model specification

disp_mod_ps_data <- displacement_data %>% 
  filter(!is.na(country),
         gear == "purse_seines") %>% 
  mutate(post = post *1,
         year2 = year^2,
         month_c = as.character(month))

disp_mod_ps1 <- lm(h_prop ~ post*country, data = disp_mod_ps_data)

models <- list(disp_mod_ps1)

stargazer::stargazer(models,
                     se = commarobust::makerobustseslist(models),
                     t.auto = T,
                     p.auto = T,
                     type = "latex",
                     omit = "month",
                     omit.stat = c("adj.rsq", "f", "ser"),
                     add.lines = list(c("Month FE", "No", "Yes", "Yes")),
                     header = F,
                     single.row = T,
                     title = "\\label{tab:disp_mod}Change in the relative allocation of fishing hours by purse seiners for each region. Asterisks indicate significance levels. Numbers in parenthesis represent heteroskedastic-robust standard errors.")

```

# Discusion

Our findings provide interesting insights into the effect that LSMPAs can have on vessel behavior and the redistribution of fishing effort. These collection of results shows that the implementation of PIPA caused treated vessels to redeuce their fishing hours, and that this effect is greater for purse seiners. Even though treated vessels fish less, their relative allocation of fishing hours increased for all other fishing grouds. This fidnings does not imply that there is more fishing effort, but rather that each region receives a greater portion of the post-PIPA fishing effort (which is lower than pre-PIPA). In this section we discuss the implications of vessel-level reductions in fishing effort and the increase in relative allocation of the remaining effort through space. We also provide plausible explanations as to why purse seiners seem to be more reactive to the spatial closure.

<!-- **We should have a discussion (over the phone and in the paper) about whether a decrease in fishing hours represents a positive or negative impact of PIPA. Without data on catch it might be hard to really know.** -->

Previous studies on insular environments suggest that vessels move to distant places, which might be translated as increased costs \citep{stevenson_2013}. Nevertheless, this study does not use counterfactuals that could help account for system- or fleet-level changes that occur through time. Others have used similar satellite-tracking systems to show that fishing effort accumulates near the edges ospatial closures, yielding greater catches \citep{murawski_2005}. Yet, these vessel tracks do not cover the pre-reserve period, making it difficult identify the contribution of spatial closures to the observed spatial distribution of fishing vessels. Recent work by \citet{elahi_2018} identified that total fishing effort in a focal region where a short-term MPA was implemented showed little change, likely indicating that fishers redistributed fishing effort to compensate for the reduction in available space. Our data is assambled in a similar way, with fishing positions before and after the implementation of PIPA and vessels grouped into treated and control groups. Our BACI design, along with our difference-in-differences analysis allows us to make causal inferences about the effect that large scale marine protected areas have on fishing effort.

The different responses observed between purse seiners and longliners might have two possible explanations. It is likely that PIPA did not contain habitat that longliners would consider optimal. Therefore, the sporadic fishing events that occurred there are of little importance to the fleet, and it is unlikely that the implementation of PIPA has an effect on them. Alternatively, the differences may be due to the nature of each fishing gear. Purse seiners are often constrained by seafloor and termocline depth, and often operate in more restricted areas \citep{kroodsma_2018}. Tuna purse seiners are known to have greater proportion of null sets (*i.e.* where the purse seine is effectively casted arround tuna, but no catch is obtained) during El Niño years, where the termocline deepens in the Eastern Pacific \citep{dreyfusleon_2015}. On the other hand, longliners may be more flexible as to where they can deploy their longlines. \citet{ortuocrespo_2018} evaluated the ecological niche of the pelagic longline fleet, and suggest that the fleet may be underutilizing the ocean, meaning that they can easily redistribute elsewhere.

<!-- # Notes and things to do -->

<div id = "refs"> </div>
